Bayesian Sparsification of $\cplx$-valued networks

### 2020-01-31 paper/paper.tex
slight rewording in the abstract and some typo fixes

\begin{abstract}
With continual miniaturization ever more applications of deep learning can be found in
embedded systems, where it is common to encounter data with natural complex domain
representation. To this end we extend Variational Dropout of \citet{kingma_variational_2015}
and \citet{kharitonov_variational_2018} to complex-valued neural networks building upon
the work of \citet{trabelsi_deep_2017}. We verify the proposed Bayesian technique and
assess the performance-compression trade-off by conducting a large-scale numerical study
of $\cplx$-valued networks on two tasks: image recognition on MNIST-like and CIFAR10
datasets and music transcription on MusicNet \citep{thickstun_learning_2017}. We reliably
replicate the state-of-the-art result for $\cplx$-valued networks on MusicNet with $50-100\times$
compression rate.
\end{abstract}

@@

In this section we study the complex variational dropout techniques discussed above
on the tasks and datasets, that were studied in prior research on complex-valued networks.
All experiments share the following settings unless specified otherwise:
\begin{itemize}
  \item ADAM optimizer with the base learning rate of $10^{-3}$
  \item learning rate the annealing schedule from \citep{trabelsi_deep_2017}
  \item global gradient clipping within an $\ell_2$-norm ball of radius $0.05$
\end{itemize}
Every experiment is replicated several times to take random effects into account, e.g.
random network initialization prior to ``dense'' stage, randomly ordered batches during
SGD. Throughout every experiment on MNIST-like datasets (sec.~\ref{sub:mnist_like_datasets},
we \textit{fixed} a random subset of the train split of each dataset, and used the test
split to gauge the final performance. Same appriach was applied to CIFAR10 (sec.~%
\ref{sub:cifar10_100} experiments, which could partly expalin the overall poor test
perfomrnace we got from the tested VGG16 model ($\real$ and $\cplx$ version).

@@

At the \textit{``dense''} stage every network is fit ``as-is'', retaining its original
non-bayesian architecture and using only the likelihood from \eqref{eq:elbo} without
variational dropout.
%
During the \textit{``sparsify''} stage we make every layer bayesian and apply $\cplx$-variational
dropout sec.~\ref{ssub:variational_dropout} (or sec.~\ref{ssub:empirical_bayes}) or its $\real$
counterpart. To be able to study the attainable sparsity levels we inject a coefficient $
  C \in (0, 1]
$ at the KL divergence term in \eqref{eq:elbo}:
\begin{equation}  \label{eq:elbo_with_coef}
    - \frac{C}N KL(q_{\theta} \| \pi_{\lambda})
    + \frac1{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,. \tag{\ref{eq:elbo}'}
\end{equation}
Higher $C$ yields increasingly more sparse layers upon termination of this stage.
%
At the \textit{``fine-tune''} stage we deduce sparsity masks from the parameters of the fit
variational approximation and train the remaining non-zero parameters similarly to the
``dense'' stage, i.e. with non-bayesian layers, but with fixed sparsity masks.

@@

Indeed, as is shown by the numerical experiments on MNIST-like datasets and CIFAR10, a
real-valued model trained on the Fourier transformed data underperforms relative to
its architectural twin, fit to on the raw color data. At the same time complex-valued
models

>>>>

\begin{abstract}
With continual miniaturization ever more applications of deep learning can be found in
embedded systems, where it is common to encounter data with natural complex domain
representation. To this end we extend Variational Dropout of \citet{kingma_variational_2015}
and \citet{kharitonov_variational_2018} to complex-valued neural networks building upon
the work of \citet{trabelsi_deep_2017}. We verify the proposed Bayesian technique and
assess the performance-compression trade-off by conducting a large-scale numerical study
of $\cplx$-valued networks on two tasks: image recognition on MNIST-like and CIFAR10
datasets and music transcription on MusicNet. We reliably replicate the state-of-the-art
result by \citet{thickstun_learning_2017} for $\cplx$-valued networks on MusicNet with
$50-100\times$ compression rate.
\end{abstract}

@@

In this section we study the complex variational dropout techniques discussed above
on the tasks and datasets, that were studied in prior research on complex-valued networks.
All experiments share the following settings unless specified otherwise:
\begin{itemize}
  \item ADAM optimizer with the base learning rate of $10^{-3}$
  \item learning rate the annealing schedule from \citep{trabelsi_deep_2017}
  \item global gradient clipping within an $\ell_2$-norm ball of radius $0.05$
\end{itemize}
Every experiment is replicated several times to take random effects into account, e.g.
random network initialization prior to ``dense'' stage, randomly ordered batches during
SGD. Throughout every experiment on MNIST-like datasets (sec.~\ref{sub:mnist_like_datasets},
we \textit{fixed} a random subset of the train split of each dataset, and used the test
split to gauge the final performance. Same approach was applied to CIFAR10 (sec.~%
\ref{sub:cifar10_100} experiments, which could partly explain the overall poor test
performance we got from the tested VGG16 model ($\real$ and $\cplx$ version).

@@

At the \textit{``dense''} stage every network is fit ``as-is'', retaining its original
non-bayesian architecture and using only the likelihood from \eqref{eq:elbo} without
variational dropout.
%
During the \textit{``sparsify''} stage we make every layer bayesian and apply $\cplx$-variational
dropout sec.~\ref{ssub:variational_dropout} (or sec.~\ref{ssub:empirical_bayes}) or its $\real$
counterpart. To be able to study the attainable sparsity levels we inject a coefficient $
  C \in (0, 1]
$ at the KL divergence term in \eqref{eq:elbo}:
\begin{equation}  \label{eq:elbo_with_coef}
    - \frac{C}N KL(q_{\theta} \| \pi_{\lambda})
    + \frac1{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,. \tag{\ref{eq:elbo}'}
\end{equation}
Higher $C$ yields increasingly more sparse layers upon termination of this stage.
%
At the \textit{``fine-tune''} stage we deduce sparsity masks from the parameters of the fit
variational approximation and train the remaining non-zero parameters similarly to the
``dense'' stage, i.e. with non-Bayesian layers, but with fixed sparsity masks.

@@

Indeed, as is shown by the numerical experiments on MNIST-like datasets and CIFAR10, a
real-valued model trained on the Fourier transformed data underperforms relative to
its architectural twin, fit to on the raw colour data. At the same time complex-valued
models
